---
title: "Report C (Group 2.A)"
author: "S2156991, S2148653, S2175591"
output: bookdown::pdf_document2
date: "2023-01-17"
toc: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE) 
```

```{r include=FALSE}
library(tidyverse)
library(patchwork)
library(kableExtra)
```


```{r include=FALSE}
pass_scores <- read_csv("https://uoepsy.github.io/data/dataset-ipeds-2012-subset2.csv")
dim(pass_scores)
head(pass_scores)
```
```{r, include=FALSE}
glimpse(pass_scores)
summary(pass_scores)
```
# Introduction
The report samples 10,725 units, based on graduation rates, the type of college and the region they are situated. The data was used to provide the average graduation rates of both male and female students respectively. The recorded five variables include, "type" -1 to 4 to represent college type. "Region" represents the regions in which the colleges are located. 
"Gradrate" represents the graduation rate of all the students. 
"Gradratem" represents graduation rate of male students. 
"Gradratew" represents graduation rate of female students. 
The data doesn't include impossible scores, with all data being within the range of 0-100. 
The variables of interest in our report will be the graduation rate of women across various colleges in the US. 

# Analysis
```{r include=TRUE, fig.height=4}
plt_hist <-ggplot(pass_scores, aes(x = gradratew)) + 
  geom_histogram(colour = 'light blue', fill = 'dark blue')

plt_box <- ggplot(pass_scores, aes(x = gradratew)) + 
    geom_boxplot(fill = 'light blue', colour = 'dark blue')

plt_hist/plt_box


```
The figures here demonstrate the female graduation The shape of distribution is a normal distribution, with no significant outliers

This table (Table1) demonstrates thee minimum and maximum graduation rates, alongside the mean graduation rate across all colleges as well as the standard deviation. The minimum graduation rate of women is 3% and the maximum is 100%. The average graduation rate for this data set is 51.3% and the standard deviation is 20.8. 

```{r}
stats <- pass_scores %>%
    summarise(n = n(),
              Min = min(gradratew),
              Max = max(gradratew),
              M = mean(gradratew),
              SD = sd(gradratew))

tab_stats <- stats %>%
    kbl(booktabs = TRUE, digits = 2, 
        caption = "Descriptive statistics for female graduation rate") %>%
  kable_styling(latex_options = "HOLD_position")
tab_stats
```

```{r include=FALSE}
# Sample mean
xbar <- stats$M

# Standard error
s <- stats$SD
n <- stats$n
se <- s / sqrt(n)
se
```
Standard error of the mean graduation rate is 0.45
```{r include =FALSE}
# Quantiles
tstar <- qt(c(0.025, 0.975), df = n - 1)
tstar
```
The quantiles of our t distribution, are -1.96 and 1.96
```{r include = FALSE}
# CI
xbar + tstar * se
```
Our 95% confidence interval is 50.45 to 52.21. From the data, we found the average graduation rate in females is 51.3%, and therefore we are 95% confident that the mean female graduation rate will fall between these intervals. 
```{r include = FALSE}
statsmin <- pass_scores %>%
    summarise(n = n(),
              Min = min(gradrate))
statsmin

```
$$H_{0} : \mu = \mu_{50}$$
The null hypothesis we are proposing is that the average graduation rate of women in US is 50%
$$H_{1} : \mu \neq \mu_{50}$$
Our alternative hypothesis suggests that the average graduation rate of women in US not equal to 50%
```{r include = FALSE}
# Sample mean
xbar <- stats$M

# Standard error
s <- stats$SD
n <- stats$n
se <- s / sqrt(n)

# Observed t-statistic
tobs <- (xbar - 50) / se

tobs

```
As the sample size is 
 n = 2145, if the null hypothesis is true the t-statistic will follow a t(2144) distribution.
 
```{r include = FALSE}
2 * pt(abs(tobs), df = n - 1, lower.tail = FALSE)
```
 At the 5% significance level, the data provides strong evidence in against the null hypothesis given that the p-value of 0.003 is less than the significance level of 0.05 in this case. Instead we accept the alternative hypothesis which states that the mean graduation for women in US is different from the rate suggested by our null hypothesis average of 50: t(2144) = 2.97, p = 0.003, two-tailed.
 


 